#---
#title: "chapter6"
#author: "Miikka Haapakoski"
#date: "12/2/2020"
#output: html_document
---

```{r}
#Libraries
library(dplyr)
library(ggplot2)
library(tidyr)
library(corrplot)
library(knitr)
```

#Easiest way of getting the data imported, is to get RATS and MABS from the last data wranling phase. 
```{r}
BPRSL <- read.table(file="~/R/win-library/4.0/IODS-project/data/BPRSL.txt", header = TRUE, sep  =" ")
RATSL <- read.table(file="~/R/win-library/4.0/IODS-project/data/RATSL.txt", header = TRUE, sep = '\t')

summary(BPRSL)
summary(RATSL)

dim(BPRSL)
dim(RATSL)


```

```{r}
# Factor variables ID and Group
RATSL$ID <- factor(RATSL$ID)
RATSL$Group <- factor(RATSL$Group)

#Table 8.1 with RATSL data

kable(RATSL)

# Plot the RATSL data
ggplot(RATSL, aes(x = Time, y = Weight, group = ID)) +
  geom_line(aes(linetype = Group)) +
  scale_x_continuous(name = "Time (days)", breaks = seq(0, 60, 10)) +
  scale_y_continuous(name = "Weight (grams)") +
  theme(legend.position = "top")

``` 
#Figure 8.1 with RATSL data 
```{r}
ggplot(RATSL, aes(x = Time, y = Weight, linetype = ID)) +
  geom_line() + scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ Group, labeller = label_both) + 
  theme_bw() + theme(legend.position = "none") +
  theme(panel.grid.minor.y = element_blank()) +
  scale_y_continuous(limits = c(min(RATSL$Weight), max(RATSL$Weight)))
```

#Figure 8.2. with RATSL data
```{r}
# Standardise the variable Weight
RATSL <- RATSL %>%
  group_by(Time) %>%
  mutate(stdweight = (Weight - mean(Weight))/sd(Weight) ) %>%
  ungroup()

# Glimpse the data
glimpse(RATSL)

# Plot again with the standardised Weight
ggplot(RATSL, aes(x = Time, y = stdweight, linetype = ID)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ Group, labeller = label_both) +
  scale_y_continuous(name = "standardised Weight") +
  theme_bw() + theme(legend.position = "none")
```
## Figure 8.3 with RATSL

```{r}
# Number of weeks, baseline (week 0) included
n <- RATSL$Time %>% unique() %>% length()

# Summary data with mean and standard error
RATSS <- RATSL %>%
  group_by(Group, Time) %>%
  summarise( mean=mean(Weight), se=sd(Weight)/sqrt(n) ) %>%
  ungroup()

# Glimpse the data
glimpse(RATSS)

# Plot the mean profiles
ggplot(RATSS, aes(x = Time, y = mean, linetype = Group, shape = Group)) +
  geom_line() + scale_linetype_manual(values = c(1,2,3)) +
  geom_point(size=3) + scale_shape_manual(values = c(1,2,3)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se, linetype="1"), width=0.3) +
  theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(legend.position = c(0.8,0.3)) +
  scale_y_continuous(name = "mean(weight) +/- se(weight)")

# Create a summary data by  and Weight with mean as the summary variable (ignoring baseline week 0).
RATSL8S <- RATSL %>%
  filter(Time > 0) %>%
  group_by(Group, ID) %>%
  summarise( mean=mean(Weight) ) %>%
  ungroup()

# Glimpse the data
glimpse(RATSL8S)

# Draw a boxplot of the mean versus Weight
ggplot(RATSL8S, aes(x = Group, y = mean)) +
  geom_boxplot() +
  stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "mean(Weight), Time")
  
```

```{r}
# Create a new data by filtering the outlier group and adjust the ggplot code the draw the plot again with the new data
RATSL8S1 <- RATSL8S %>%
  filter(Group != 3)
  
# Glimpse the data
glimpse(RATSL8S1)

  
# Perform a two-sample t-test with groups 1 and 2


t.test(mean ~ Group, data = RATSL8S1, var.equal = TRUE)
kable(RATSL8S1)
```

#Table and figure 9.1 with BPRSL 

# Factor treatment & subject
BPRS$treatment <- factor(BPRS$treatment)
BPRS$subject <- factor(BPRS$subject)

```{r}
kable(BPRSL)

ggplot(BPRSL, aes(x = week, y = bprs, group = subject)) +
  geom_text(aes(label = treatment)) +
  scale_x_continuous(name = "Time (weeks)", breaks = seq(0, 8, 1)) +
  scale_y_continuous(name = "bprs (score)") + 
  theme(legend.position = "top")

```

#Regression model 

```{r}
# create a regression model BPRSL_reg
BPRSL_reg <- lm(bprs ~ week + treatment, data = BPRSL)

# print out a summary of the model
summary(BPRSL_reg)
```

```{r}
# access library lme4
library(lme4)

# Create a random intercept model
BPRSL_ref <- lmer(bprs ~ week + treatment + (1 | subject), data = BPRSL, REML = FALSE)

# Print the summary of the model
summary(BPRSL_ref)


# create a random intercept and random slope model
BPRSL_ref1 <- lmer(bprs ~ week + treatment + (week | subject), data = BPRSL, REML = FALSE)

# print a summary of the model
summary(BPRSL_ref1)

# perform an ANOVA test on the two models
anova(BPRSL_ref1, BPRSL_ref)

```
```{r}
# create a random intercept and random slope model
BPRSL_ref2 <- lmer(bprs ~ week * treatment + (week | subject), data = BPRSL, REML = FALSE)
# print a summary of the model
summary(BPRSL_ref2)
# perform an ANOVA test on the two models
anova(BPRSL_ref2, BPRSL_ref1)
```

