#---
#title: "chapter5.Rmd"
#author: "Miikka Haapakoski"
#date: "11/26/2020"
#output: html_document
#---
```{r}
library(dplyr)
library(ggplot2)
library(GGally)
library(tidyr)
library(openxlsx)
library(corrplot)
library(knitr)
```

#Analysis. Show a graphical overview of the data and show summaries of the variables in the data. Describe and interpret the outputs, commenting on the distributions of the variables and the relationships between them. (0-3 points)


#1. Summary part 
```{r}
human <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human2.txt", sep=",", header=TRUE)

dim(human)
str(human)
summary(human)


ggpairs(human)

cor(human) %>% corrplot
```

#2. 
```{r}
# standardize the variables
human_std <- scale(human)

summary(human_std)

# perform principal component analysis (with the SVD method)
pca_human <- prcomp(human_std)

# draw a biplot of the principal component representation and the original variables
biplot(pca_human, choices = 1:2, cex = c(0.8, 1), col = c("grey40", "deeppink2"))
```

#3. 
```{r}
# create and print out a summary of pca_human
s <- summary(pca_human)
s

# rounded percentages of variance captured by each PC
pca_pr <- round(100*s$importance[2,], digits = 1) 

# print out the percentages of variance
pca_pr

# create object pc_lab to be used as axis labels
pc_lab <- paste0(names(pca_pr), " (", pca_pr, "%)")

# draw a biplot
biplot(pca_human, cex = c(0.8, 1), col = c("grey40", "deeppink2"), xlab = pc_lab[1], ylab = pc_lab[2])
```
#5. 
# install and access the FactoMineR package
```{r}
if (!require("devtools")) install.packages("devtools")
library(devtools)
install_github("husson/FactoMineR")

library (FactoMineR)

# load the data
data("tea")

# explore the dataset
str(tea)
dim (tea)

# column names to keep in the dataset
keep_columns <- c("Tea", "How", "how", "sugar", "where", "lunch")

tea_time <- select(tea, one_of(keep_columns))


# visualize the dataset
gather(tea_time) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar() + theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))

# multiple correspondence analysis
mca <- MCA(tea_time, graph = FALSE)

# summary of the model
summary(mca)

# visualize MCA
plot(mca, invisible=c("ind"), habillage = "quali")

#According to the plots, most of the people drink tea from the tea bags without anything added. Tea is mainly drank not on the lunch and sugar is about 50%. Earl grey is the most popular tea and its bought from the chain store. 
